---
title: "Lecture 5"
author: "Zhenya Kosovan"
date: "2019/10/25"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

## Structure
- Import & Export
- Aggregation & Mutation
```{r setup, include = FALSE}
library(tidyverse)
```

---
class: middle, inverse

# Import & Export

---
class: middle, inverse

# Mutation & Aggregation

---

## Mutation
- adding a new or changing existing columns
```{r mutate}
econ =
    readxl::read_excel("data/econmap.xlsx", 1) %>%
    select(name, year, gdp_cap, energy_productivity, population, labor_force, tfp)
head(econ)
```

---

## Mutation in baseR
- ading new or changing existing columns
```{r mutate_base}
econ$region = ifelse(econ$name == "Afghanistan", "Afghan", "Others") # base R
head(econ)
econ$region = NULL # erase column
colnames(econ)

```

---

## Mutation in Tidyverse
- ading new or changing existing columns

```{r mutate_tidy}
econ %>%
    mutate(region = ifelse(name == "Afghanistan", "Afghan", "Others"),
           tfp = NULL)
```

```{r region, include = FALSE}
econ =
    econ %>%
    mutate(region = ifelse(name == "Afghanistan", "Afghan", "Others"),
           tfp = NULL)
```

---

## Aggregation
- calculate function by group

```{r aggragate}
econ %>% 
    group_by(name) %>% 
    summarise(gdp = mean(gdp_cap),
              energy = sum(energy_productivity)
              )
```

---

## Aggregation on multi-index

```{r aggregate_multi}
econ %>% 
    group_by(region, year) %>% 
    summarise(pop = mean(population),
              labor = mean(labor_force)) %>% 
    ungroup() %>% 
    mutate(active = labor / pop) 
```