---
title: "Lecture 10"
author: "Zhenya Kosovan"
date: "2019/11/29"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

## Structure
- ggplot2 ideology
- Practice

```{r setup, include = FALSE}
library(tidyverse)
```

---
class: middle, inverse

# ggplot2

---

## Why ggplot2?
- consistent underlying grammar of graphics
- plot specification at high level of abstraction
- very flexible
- theme system polishing experience
- mature and complete system

Some limitations (yeah, there are some):

-   3-dimensional graphics 
-   Graph-theory type graphs
-   Interactive graphics

---

## Grammar of graphics
Idea: construct a plot out of independent building blocks:
- data
- aesthetic mapping
- geometric object
- statistical transformations
- scales
- coordinate system
- position adjustments
- faceting


---
## Introducing the data

```{r}
housing <- read_csv("dataSets/landdata-states.csv")

```

---

## Simple functions

```{r}
head(housing[1:5])
```

---

## Basic Histogram

```{r fig1, fig.height = 5, fig.width = 8}
library(ggplot2)
ggplot(housing, aes(x = Home.Value)) +
  geom_histogram()

```


---

## Basic Scatterplot

```{r, fig.height = 5, fig.width = 8 }
ggplot(filter(housing, State %in% c("MA", "TX")),
       aes(x=Date,
           y=Home.Value,
           color=State))+
  geom_point()

```

---

# Geometric Objects And Aesthetics

## Aesthetic Mapping

In ggplot land **aesthetic** means "something you can see", i.e.:

-   position (i.e., on the x and y axes)
-   color ("outside" color)
-   fill ("inside" color)
-   shape (of points)
-   linetype
-   size

Aesthetics is different for different types of plots:
- **geom help** is there for assistance
- aesthetic mappings are set with the **`aes()`** function.

---
## Geometic Objects (`geom`)

Geometric objects are the actual marks we put on a plot. Examples include:

-   points (`geom_point`, for scatter plots, dot plots, etc)
-   lines (`geom_line`, for time series, trend lines, etc)
-   boxplot (`geom_boxplot`, for, well, boxplots!)

A plot must have at least one geom; there is no upper limit. You can add a geom to a plot using the `+` operator

You can get a list of available geometric objects using the code below:

```{r}
help.search("geom_", package = "ggplot2")
```



---
```{r,fig.height = 6, fig.width = 8}
hp2001Q1 <- filter(housing, Date == 2001.25) 
ggplot(hp2001Q1,
       aes(y = Structure.Cost, x = log(Land.Value))) +
  geom_point()
```


---

```{r,fig.height = 6, fig.width = 8}
hp2001Q1$pred.SC <- predict(lm(Structure.Cost ~ log(Land.Value),
                               data = hp2001Q1))

p1 <- ggplot(hp2001Q1, aes(x = log(Land.Value),
                           y = Structure.Cost))

p1 + geom_point(aes(color = Home.Value)) +
  geom_line(aes(y = pred.SC))
```

---
## Smoothing

```{r,fig.height = 4, fig.width = 8}
p1 +
  geom_point(aes(color = Home.Value)) +
  geom_smooth()
```

---
## Text points

```{r,fig.height = 6, fig.width = 8}
p1 + 
  geom_text(aes(label=State), size = 3)
```



---
## Text points improved

```{r,fig.height = 5, fig.width = 8}
## install.packages("ggrepel") 
library("ggrepel")
p1 + 
  geom_point() + 
  geom_text_repel(aes(label=State), size = 3)
```


---
## Utilizing Extra Aesthetics Options

```{r,fig.height = 5, fig.width = 8}

p1 +
  geom_point(aes(color=Home.Value, shape = region))
```

---
## Scaling'n'Stuff
- if a variable is in *aes()* it is mapped, i.e. aes(color = x) doesn't specify color
- but you can manipulate how exactly!
- you do that by modifying **scale**

In ggplot2 possible scales are:
- position
- color and fill
- size
- shape
- line type

To change scale you need need to add **scale_<aestheric>_<type>** to the plot

---
## Scaling Example
```{r,fig.height = 3, fig.width = 8}
p2 <- ggplot(housing,
             aes(x = State,
                 y = Home.Price.Index)) + 
        theme(legend.position="top",
              axis.text=element_text(size = 6))
(p3 <- p2 + geom_point(aes(color = Date),
                       alpha = 0.5,
                       size = 1.5,
                       position = position_jitter(width = 0.25,
                                                  height = 0)))
```

--- 
## Scaling Example

```{r,fig.height = 3, fig.width = 8}
p3 + scale_x_discrete(name="State Abbreviation") +
  scale_color_continuous(name="",
                         breaks = c(1976, 1994, 2013),
                         labels = c("'76", "'94", "'13"))
```

---
## Scaling Example

```{r,fig.height = 5, fig.width = 8}
p3 +
  scale_x_discrete(name="State Abbreviation") +
  scale_color_continuous(name="",
                         breaks = c(1976, 1994, 2013),
                         labels = c("'76", "'94", "'13"),
                         low = "blue", high = "red")
```

---
## Even More Scaling!
```{r,fig.height = 4.5, fig.width = 8}
p3 +
  scale_color_gradient2(name="",
                        breaks = c(1976, 1994, 2013),
                        labels = c("'76", "'94", "'13"),
                        low = "blue",
                        high = "red",
                        mid = "gray60",
                        midpoint = 1994)
```

---

## Faceting
- many lines on one graph **are not** representative
- faceting -> splitting all lines into small separate graphs with common aesthetics
- makes comparison between the plots easier

Faceting is implemented by **facet_wrap()** and **facet_grid()**

---
```{r,fig.height = 5, fig.width = 8}
p4 = ggplot(housing, aes(x = Date, y = Home.Value))
p4 + geom_line(aes(color = State))  
```

---
```{r,fig.height = 7, fig.width = 8}
(p4 = p4 + geom_line() +
   facet_wrap(~State, ncol = 10))
```

---
## Themes aka handling non-data elements
- axis labels
- plot background
- facet label background
- legen appearance

---
```{r,fig.height = 8, fig.width = 8}
p4 + theme_linedraw()
```

---
```{r,fig.height = 7, fig.width = 8}
p4 + theme_minimal() +
  theme(text = element_text(color = "turquoise"))
```