---
title: "Lecture 2"
author: "Zhenya Kosovan"
date: "2018/10/03"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

## Structure
- Data & Variable types
- Arithmetic & logic operators
- Subsetting (slicing)

---
class: middle, inverse

# Data & Variable types

---

## Variables
--

- any text that is not enclosed in "" or ''
- without brackets at the end, e.g. test()
- not predefined words

```{r variables}
"number"
number = 5
number
"text"
text <- 'R is amazing'
print(text)
```

---

## Data types
--

```{r data_types, eval = F}
input
"text"
1.5
20L
TRUE
FALSE
Sys.Date()
Sys.time()
0 / 0
```

---
## Data types cont.

```{r data_types_comments, eval = F}
input # variable (not an actual data type)
"text" # character
1.5 # numeric
20L # integer
TRUE #boolean
```
```{r data_types_comments_eval}
Sys.Date() # Date
Sys.time() # POSIXct POSIXlt
NA; 0 / 0 # this is bad idea
```

---

## Casting
- how to change data types?

```{r casting}
as.character(1)
as.numeric("1")
as.integer(1.5)
```

---
## Data structures
- what a variable can be?



--
- vector

--
- matrix

--
- list

--
- data frame

---

## Vectors

```{r vector}
c(2, 5)
c(8, 25, 3.14, -4)
c('R', 'S')
c(25, T, 'a') # T is the same as TRUE
2:6
```

---

## Factors - special data type

```{r factors}
fac = factor(c('a','a','b','c'), levels = c("a","b","c"))
fac
as.numeric(fac)
```

---

## Matrices

```{r matrix}
matrix(c(2, 5))
matrix(c(8, 25, 3.14, -4), nrow = 2)
matrix(c('R', 'S', 'C', 'Py', 'C', 'J'), ncol = 3, byrow = TRUE)
```

---

## Matrices cont.

```{r matrix_cont}
matrix(c('R', 5, 'C', NA, Sys.Date() , T), nrow = 3, ncol = 2)
```

---

## Data frames

```{r data_frame}
input = data.frame(wage = c(1:3), gender = c("male", "female", NA))
input
rownames(input)
colnames(input)
```

---

## Data frames cont.

```{r data_frames_question}
output = data.frame(wage = c(1 , 2, NA, Sys.time(), 'not avaliable'))
```

--

```{r data_frames_answer}
output
```

---

## Lists

```{r list}
list(125, "R is good")
list(onetyone = 11, zwoelf = "12")
list(a = c(1:5))
```

---

## Lists cont.

```{r lists_cont}
list(WT = data.frame(NUM = 1:2, STR = c('a', 'b')),
     OM = matrix(1:12, ncol = 4))
```


# Arithmetic & logic operators

---

## Arithmetic operators

```{r arithmetics}
2 + 7
x = 2; y = 7
x + y
y - x
y^x
y/x
```

---

## Arithmetics operators cont

```{r arithmetics_cont}
x = 2; y = 7
y %% x
y %/% x
```

---

## Arithmetic functions

```{r arithmetics_functions}
x = 1:10
sqrt(x) # log(), exp(), sin(), ...
sum(x)
mean(x) # median(), quantile(), ...
```

---

## Arithmetic functions cont.

```{r arithmetics_functions_cont}
summary(x) # can be used on data.frame and many other types 
```

---

## Sequences

```{r sequences}
x = 6; y = 15
x:y
seq(x, y, by = 2.5)
seq_len(y)
seq_along(x:y)
rep(y, x)
```

---

## Logic operators

```{r boolean}
TRUE & FALSE
c(T, F) & T
c(T, F) && T
T | F
!F
```

---

## Comparison

```{r comparison}
"a" == "b"
2 != 5
2 %in% 1:5
3.985 < 7
5.61 >= 5.61
```


---
class: middle, inverse

# Subsetting (slicing)

---

## Names (indexing)
- vectors, lists can have position names
- data.frames and matrices can have rownames and colnames

```{r naming}
c(a = 3, "b" = 4)

namevec = 1:3
names(namevec) = c('a', 'b', 'c'); namevec

list(g = 1, "h" = "R is very kind")
```

---

## Names cont.

```{r naming_cont}
namemat = matrix(1:4, 2)
colnames(namemat) = c("eq.1","eq.2")
rownames(namemat) = c("x","y")
namemat

namedf = data.frame("a" = 1L:3L, b = c('a','b','c'))
rownames(namedf) = c("Jana","Mirka","Lisa")
namedf
```

---

## Slicing in general
- can be done via three specifications
  - numeric position, e.g. third, fifth and seventh position
  - names, e.g. position 'Jana' and 'Mirka'
  - boolean, e.g. TRUE, FALSE, TRUE (same as the first and third position)

---

### Slicing vectors

```{r slice_vec}
testvec = 1:6
names(testvec) = c("a", 'b', "c", "d", 'e', "f")
testvec[2]
testvec[c(3,5)]
testvec[1:3]
testvec[c('a','d')]
```

---

### Slicing vectors cont.

```{r slice_vec_cont}
testvec[testvec < 4]
sec_vec = 6:1
testvec[sec_vec < 4]
```

---

### Slicing lists

```{r slice_list}
testlist = list(a = matrix(2,1), "b" = c("R","is very kind"), c = T, d = NA)
testlist[c(2, 4)]
testlist[c('c','a')]
```

---

### Slicing lists cont.

```{r slice_list_cont}
testlist[1]
testlist[[1]]
# alternatively
testlist$a

testlist[[2]][1]; testlist$b[1]
```

---

### Slicing matrices

```{r slice_matrix}
namemat

# no dollar notation
# names hard to use
namemat[2,c(1, 2)]
namemat[c(1, 2), 2]
```

---

### Slicing data.frames

```{r slice_df}
namedf
namedf[1,c(1,2)]
namedf[2,'a']
namedf['Jana','b']
```

---

### Slicing data.frames cont.

```{r slice_df_cont}
namedf[c('Jana','Mirka'), c('a','b')]
namedf$a
namedf$b[1]
namedf[namedf$a < 2,]
```

